Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-17 paper_jasa.tex
proof-read through page 14

\green{Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in deciding on the choice of plot to use.} Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose. This is akin to different choices of statistics in conventional hypothesis testing, for example, mean, median, or trimmed mean. One is typically better than another. For two different visual test statistics of the same actual data, one is considered to be better, if $T(y)$ is more easily distinguishable to the observer. Power is typically used to measure this characteristic of a test. %Power is the probability of rejecting $H_0$ when it is not true.  We can assess power therefore both empirically based on experimental data and through theoretical considerations.

@@

Suppose each of $K$ independent observers gives evaluations on multiple lineups, and responses are considered to be binary random variable, $X_{\ell i}$, where $X_{\ell i} = 1$, if subject $i$ correctly identifies the actual data plot on lineup $\ell$, $\ell = 1, ..., L$, and 0 otherwise. A mixed effects logistic regression model is used for $p_{\ell i} = E(X_{\ell i})$, accommodating both for different abilities of observers as well as differences in the difficulty of lineups.

Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the actual data plot from lineup $\ell$, and 0 otherwise.
The model can be  fit as:% $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
\begin{equation} \label{eqn:mixed}
g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
\end{equation}

@@

\begin{tabular}{lp{5.5in}}
$W$  & is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$. Covariates could include  demographic information of individuals, such as age, gender, education level etc. as well lineup-specific elements, e.g. effect size or difficulty level. \\
$\delta$ & is a vector of parameters for fixed effects.\\
$Z_{\ell i}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
$\tau$ & is a vector of independent real-valued random variables $\tau_{\ell i}$,  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$. $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups. 
Note that $t_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.\\
\end{tabular}

\noindent Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leands to the estimate of the subject and the lineup specific probability of successful evaluation as 
\begin{equation} \label{eqn:mixed_power}
\hat p_{\ell i} =  g^{-1}(W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}).
\end{equation}

@@

The following properties can only be derived for the situation of a fully parameterized  simulation study, as conducted in this paper. They allow for a direct comparison of lineup tests against the conventional counterparts, and also to identify properties relevant for a quality assessment of lineups, for when they are used in practical settings. Two assumptions are critical:
\begin{enumerate} \itemsep 0in
\item  the plot setup is structured in a way that makes it possible for an observer to identify a deviation from the null hypothesis,
\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
\end{enumerate}
Evidence in support of the second assumption will be seen in the data from the study discussed in Section~\ref{sec:simulation}, the degree to which the first assumption is fulfilled is reflected by the power of a lineup. The better suited a design is for a particular task, the higher its power will be.

@@

%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the conventional test statistic.
%, with non-centrality parameter $\delta$. 
\noindent Then the distribution function of the $p$-value, $p_D$, under $H_0$, is uniform, since:  
\begin{eqnarray}\label{dist_p}
\green{F_{p_D}}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
  &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
  &=& 1-Pr\left(|t_{obs}| \le F_{|t|}^{-1}(1-p)\right) \nonumber \\
  &=&%\left\{ \begin{array}{ll}
          1-F_{|t|}\left( F_{|t|}^{-1}(1-p)\right)=p \mbox{ ; under $H_0$} 
 %         1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
%       \end{array} \right.     
\end{eqnarray}

@@

\green{The above lemma allows two immediate conclusions for the use of lineups.
The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$, since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the actual data plot decreases with $m$. This may seem counterintuitive, but it makes sense. With fewer null plots the actual data plot will stand out, when the alternative hypothesis is true. Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$ -- if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and less affected by an increase in $m$. This can also be seen in Figure \ref{fig:pval_power}, which illustrates lemma \ref{lemma}.   Figure \ref{fig:pval_power} shows the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance. It can also be seen that for a $p$-value, $p_D$, of about 0.15 for the data plot, the signal in the plot is so weak, that it cannot be distinguished from null plots in a lineup of size $m=20$.} % This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
%also seen in our experiments as shown in Figure \ref{fig:pval_pcorrect}.

@@

In this setting, there are many established graphics that are used to evaluate and diagnose the fit of a regression model (e.g. \citet{cook:99}). Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and the visual test statistic that might be used for testing these hypotheses. These plots should be familiar, because they are all in common use. For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ (cases 1-4). To assess whether the linear assumption is appropriate we would make a plot of residuals against fitted values (case 5). For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and catgorical explanatory variable respectively.

@@

Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$. If the responses for the two levels of the categorical variable $X_k$ in the model are different and we fit the null model to the actual data, the resulting residuals should show a significant difference between the two groups. To test this boxplots of the residuals conditioned on the two levels of $X_k$ are made, as the observed test statistic. If $\beta_k\ne 0$ the boxplots should show a vertical displacement.

@@

The conventional test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.  It forms the benchmark upon which we evaluate the visual test. To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.  This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the conventional $p$-value as details given in Lemma \ref{lemma}. Thus the expected power of the visual test in this scenario is

@@

Participants for all the experiments were recruited through \cite{turk} Amazon's Mechanical Turk. A summary of the data obtained for all three experiments are shown in Table \ref{tbl:summary}. Participants were asked to select the plot they think best matched the question given, provide a reason for their choice, and say how confident they are in their choice. Gender, age, education and geographic location of each participant are also collected. The data was collected through the web. For each of the experiments one of the lineups was used as a test plot (easy plot) which everyone should get correct, so that a measure of the quality of the subjects effort could be made.

@@

The experiment is designed to study the ability of human subjects to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}). Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical conventional test. Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian. Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}. Three replicates at each level were generated. These produced 60 different ``actual data sets'', and thus, 60 different lineups. To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates. The actual data plot was generated by making a side-by-side boxplot of the residuals (Table \ref{tbl:stat_multiple}, case 3). The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and these were plotted in the same way as side-by-side boxplots. The actual data plot was randomly placed among these null data plots to produce the lineup. Figure \ref{fig:test_category} is an example of one of these lineups. The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.

@@

The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power. Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}. Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$. For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05. Thus we have $ME =1.96 \sqrt{ \gamma(1-\gamma) / n_{\gamma} } \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$

Each subject viewed at least 10 lineups and we allowed them to provide more evaluations if they wanted. Depending on the parameter combinations we group the lineups in different difficulty levels such as easy, medium, hard and mixed (actual numbers are given in the supplementary material). For each difficulty level a specific number of lineups were randomly picked to show for evaluation. This number is chosen so that total number of evaluations for each lineup for that group exceed the threshold $n_{\gamma}$. To satisfy this plan we needed to recruit at least 300 subjects.

@@

The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid. To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data. 
An example lineup from this experiment is shown in Figure \ref{fig:test_continuous}.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11. This lineup, we think, is a difficult one. To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters. For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest). For one combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups. Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0. The order that the lineups are shown to participants is randomized. Each subject saw a substantial number of difficult lineups. \hh{While this allows us a better estimation for the power in small slopes, }
%The intention was noble, to get a better handle on the power of the lineups for small slopes, 
but practically this made the study very tiring for the participants. %It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. 
\green{perhaps should go into so much detail about this, or keep this for the conclusions section.}

@@

The first two simulation experiments use data generated with a normal error model, satisfying the conditions for conventional test procedures. In these situations there exists a test, and there would, in general, be no need to use visual inference. The simulation is conducted with the hopes that the visual test procedure, although not expected to beat the conventional test, will compare favorably with it. This third simulation is closer to the mark for the purpose of visual inference. The assumptions for the conventional test are violated, by contaminating the data. The contamination makes the estimated slopes effectively 0, yet the true value of slope parameter is not.  The data is generated from the following model:

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=n+1, ...,n*c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$, $c$ is the proportion of contaminated data. For the experiment we consider $n=100$, $c=0.15$ producing actual data with $115$ points. Also $\alpha=0$ and $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$. %Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  
A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.

>>>>

Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in deciding on the choice of plot to use. Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose. This is akin to different choices of statistics in conventional hypothesis testing, for example, mean, median, or trimmed mean. One is typically better than another. For two different visual test statistics of the same actual data, one is considered to be better, if $T(y)$ is more easily distinguishable to the observer. Power is typically used to measure this characteristic of a test. %Power is the probability of rejecting $H_0$ when it is not true.  We can assess power therefore both empirically based on experimental data and through theoretical considerations.

@@

Suppose each of $K$ independent observers gives evaluations on multiple lineups, and responses are considered to be binary random variable, $X_{\ell i} \sim  Binom_{1, p_{\ell i}}$, where $X_{\ell i} = 1$, if subject $i$ correctly identifies the actual data plot on lineup $\ell$, $1 \le \ell \le L$, and 0 otherwise. A mixed effects logistic regression model is used for $P(X_{\ell i} = 1) =  p_{\ell i} = E(X_{\ell i})$, accommodating both for different abilities of observers as well as differences in the difficulty of lineups.

%Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the actual data plot from lineup $\ell$, and 0 otherwise.
The model can be  fit as:% $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
\begin{equation} \label{eqn:mixed}
g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
\end{equation}

@@

\begin{tabular}{lp{5.5in}}
$W$  & is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$. Covariates could include  demographic information of individuals, such as age, gender, education level etc. as well lineup-specific elements, e.g. effect size or difficulty level. \\
$\delta$ & is a vector of parameters for fixed effects.\\
$Z_{\ell i}$ & $1 \le i \le K$, $1 \le \ell \le L$  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$.  \\
$\tau$ & is a vector of independent normally distributed random variables $\tau_{\ell i}$, with  $\tau  \sim  MVN(0,\sigma_\tau I_{KL \times KL})$. $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups. 
Note that $t_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.\\
\end{tabular}

\noindent Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leads to the estimate of the subject and the lineup specific probability of successful evaluation as 
\begin{equation} \label{eqn:mixed_power}
\hat p_{\ell i} =  g^{-1}(W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}).
\end{equation}

@@

The following properties can only be derived for the situation of a fully parameterized  simulation study, as conducted in this paper. They allow for a direct comparison of lineup tests against the conventional counterparts, and also to identify properties relevant for a quality assessment of lineups when they are used in practical settings. Two assumptions are critical:
\begin{enumerate} \itemsep 0in
\item  the plot setup is structured in a way that makes it possible for an observer to identify a deviation from the null hypothesis,
\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
\end{enumerate}
Evidence in support of the second assumption will be seen in the data from the study discussed in Section~\ref{sec:simulation}, the degree to which the first assumption is fulfilled is reflected by the power of a lineup. The better suited a design is for a particular task, the higher its power will be.

@@

%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the conventional test statistic.
%, with non-centrality parameter $\delta$. 
\noindent Then the distribution function of the $p$-value, $p_D$, under $H_0$, is uniform, since:  
\begin{eqnarray}\label{dist_p}
F_{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
  &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
  &=& 1-Pr\left(|t_{obs}| \le F_{|t|}^{-1}(1-p)\right) \nonumber \\
  &=&%\left\{ \begin{array}{ll}
          1-F_{|t|}\left( F_{|t|}^{-1}(1-p)\right)=p \mbox{ ; under $H_0$} 
 %         1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
%       \end{array} \right.     
\end{eqnarray}

@@

The above lemma allows two immediate conclusions for the use of lineups.
The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$, since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the actual data plot decreases with $m$. %This may seem counterintuitive, but it makes sense. 
%With fewer null plots the actual data plot will stand out, when the alternative hypothesis is true. 
Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$ -- if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and less affected by an increase in $m$. This can also be seen in Figure \ref{fig:pval_power}, which illustrates lemma \ref{lemma}.   Figure \ref{fig:pval_power} shows the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance. It can also be seen that for a $p$-value, $p_D$, of about 0.15 for the data plot, the signal in the plot is so weak, that it cannot be distinguished from null plots in a lineup of size $m=20$. % This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
%also seen in our experiments as shown in Figure \ref{fig:pval_pcorrect}.

@@

In this setting, there are many established graphics that are used to evaluate and diagnose the fit of a regression model (e.g. \citet{cook:99}). Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and commonly used statistical plots that might be used as corresponding visual test statistics.
%These plots should be familiar, because they are all in common use. 
For example, to examine the effect of variable $X_j$ on $Y$, we would plot $Y$ against $X_j$ (cases 1-4). To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5). For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable respectively.

@@

Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$. If the responses for the two levels of the categorical variable $X_k$ in the model are different, the  residuals from fitting the null model should show a significant difference between the two groups. For a visual test, we draw boxplots of the residuals conditioned on the two levels of $X_k$. If $\beta_k\ne 0$ the boxplots should show a vertical displacement.

@@

The conventional test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis, if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.  It forms the benchmark upon which we evaluate the visual test. To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.  This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the conventional $p$-value as details given in Lemma \ref{lemma}. Thus the expected power of the visual test in this scenario is

@@

Participants for all the experiments were recruited through Amazon's online web service, Mechanical Turk  \citep{turk}. A summary of the data obtained for all three experiments are shown in Table \ref{tbl:summary}. Participants were asked to select the plot they think best matched the question given, provide a reason for their choice, and say how confident they are in their choice. Gender, age, education and geographic location of each participant are also collected. %The data was collected through the web. 
For each of the experiments one of the lineups was used as a test plot (easy plot) which everyone should get correct, so that a measure of the quality of the subjects effort could be made.

@@

The experiment is designed to study the ability of human subjects to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}). Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical conventional test. Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian. Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}. Three replicates at each level were generated. These produced 60 different ``actual data sets'', and thus, 60 different lineups. To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates. The actual data plot was generated by making  side-by-side boxplots of the residuals (Table \ref{tbl:stat_multiple}, case 3). The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and  plotted in the same way. The actual data plot was randomly placed among these null data plots to produce the lineup. Figure \ref{fig:test_category} is an example of one of these lineups. The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.

@@

The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power. Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}. Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$. 
For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05. Thus we have $ME =1.96 \sqrt{ \gamma(1-\gamma) / n_{\gamma} } \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$  
% \hh{we're not even close to $n_\gamma$ with ten lineups .... we need more in the order of 400. Let's not emphasize this so much.}

Each subject viewed at least 10 lineups with the option to evaluate more.
%and we allowed them to provide more evaluations if they wanted. 
Depending on the parameter combinations we group the lineups in different difficulty levels  as easy, medium, hard and mixed (actual numbers are given in the supplementary material). For each difficulty level a specific number of lineups was randomly picked  for evaluation. This number is chosen so that total number of evaluations for each lineup for that group exceed the threshold $n_{\gamma}$. To satisfy this plan we needed to recruit at least 300 subjects.

@@

The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid. To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data. 
To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters. For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest). 
Figure \ref{fig:test_continuous} shows an example lineup for this experiment from difficulty level 4.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11.

For each combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups. Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0. The order that the lineups are shown to participants is randomized. Each subject saw a substantial number of difficult lineups.

%\hh{While this allows us a better estimation for the power in small slopes, }
%%The intention was noble, to get a better handle on the power of the lineups for small slopes, 
%but practically this made the study very tiring for the participants. %It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. 
%\green{perhaps should go into so much detail about this, or keep this for the conclusions section.}

@@

The first two simulation experiments use data generated under a normal error model, satisfying the conditions for conventional test procedures. In these situations there exists a test, and there would, in general, be no need to use visual inference. The simulation is conducted with the hopes that the visual test procedure,  will at least compare favorably with the conventional test -- without any ambition of performing equally well.
 This third simulation is closer to the mark for the purpose of visual inference. The assumptions for the conventional test are violated, by contaminating the data. The contamination makes the estimated slopes effectively 0, yet the true value of slope parameter is not.  The data is generated from the following model:

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=1, ...,n_c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$. $n_c$ is the size of thecontaminated data. For the experiment we consider $n=100$ and $n_c=15$ producing actual data with $115$ points. Further, $\alpha=0$, $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$. %Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  
A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment also follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null data plots are scatterplots of null data generated by simulating residuals from $N(0, {\hat{\sigma}}^2$) and plotting against $X_1$.
